---
title: "EDLD 652: Final Project - Chen Su"
format: 
  dashboard:
    theme: vapor
    orientation: columns
output-dir: docs
---

```{r}
library(viridis)
library(RColorBrewer)
library(colorspace)
library(ggrepel)
library(tidyverse)
library(gghighlight)
library(plotly)
library(forcats)

df <- readRDS("strata1.rds")

df <- df |>
  select(1:36) |>
  filter(!is.na(gender_coded)) |>
  filter(!is.na(income))

df <- df |>
  mutate(race_strata = factor(st_race,
                      levels = c(1, 2, 3, 4),
                      labels = c("Non-Hispanic White",
                                 "Non-Hispanic POC",
                                 "Hispanic White",
                                 "Hispanic POC"))) |> 
  relocate(race_strata, .before = gender_coded)

#data-viz 4

df <- df |>
  mutate(strata = st_race + st_gender + st_income) |>
  relocate(strata, .after = subID)

df$strata <- as.factor(df$strata)

# recode strata id to names for final data viz 4
df <- df |>
  mutate(strata_name = strata) |>
  relocate(strata_name, .after = subID)

df$strata_name <- fct_recode(df$strata_name, 
                             "High Income, Cisgender Man, Non-Hispanic White" = "111", 
                             "High Income, Cisgender Man, Non-Hispanic POC" = "112", 
                             "High Income, Cisgender Man, Hispanic White" = "113", 
                             "High Income, Cisgender Man, Hispanic POC" = "114",
                             "High Income, Cisgender Woman, Non-Hispanic White" = "121", 
                             "High Income, Cisgender Woman, Non-Hispanic POC" = "122", 
                             "High Income, Cisgender Woman, Hispanic White" = "123", 
                             "High Income, Cisgender Woman, Hispanic POC" = "124",
                              "High Income, Gender Diverse, Non-Hispanic White" = "131", 
                             "High Income, Gender Diverse, Non-Hispanic POC" = "132", 
                             "High Income, Gender Diverse, Hispanic White" = "133", 
                             "High Income, Gender Diverse, Hispanic POC" = "134",
                             "Middle Income, Cisgender Man, Non-Hispanic White" = "211", 
                             "Middle Income, Cisgender Man, Non-Hispanic POC" = "212", 
                             "Middle Income, Cisgender Man, Hispanic White" = "213", 
                             "Middle Income, Cisgender Man, Hispanic POC" = "214",
                             "Middle Income, Cisgender Woman, Non-Hispanic White" = "221",
                             "Middle Income, Cisgender Woman, Non-Hispanic POC" = "222", 
                             "Middle Income, Cisgender Woman, Hispanic White" = "223", 
                             "Middle Income, Cisgender Woman, Hispanic POC" = "224",
                              "Middle Income, Gender Diverse, Non-Hispanic White" = "231",
                             "Middle Income, Gender Diverse, Non-Hispanic POC" = "232", 
                             "Middle Income, Gender Diverse, Hispanic White" = "233", 
                             "Middle Income, Gender Diverse, Hispanic POC" = "234",
                              "Low Income, Cisgender Man, Non-Hispanic White" = "311", 
                             "Low Income, Cisgender Man, Non-Hispanic POC" = "312", 
                             "Low Income, Cisgender Man, Hispanic White" = "313", 
                             "Low Income, Cisgender Man, Hispanic POC" = "314",
                             "Low Income, Cisgender Woman, Non-Hispanic White" = "321",
                             "Low Income, Cisgender Woman, Non-Hispanic POC" = "322", 
                             "Low Income, Cisgender Woman, Hispanic White" = "323", 
                             "Low Income, Cisgender Woman, Hispanic POC" = "324",
                              "Low Income, Gender Diverse, Non-Hispanic White" = "331",
                             "Low Income, Gender Diverse, Non-Hispanic POC" = "332", 
                             "Low Income, Gender Diverse, Hispanic White" = "333", 
                             "Low Income, Gender Diverse, Hispanic POC" = "334")


```

Exploring the Variations of Perceived Neighborhood Safety across Intersectional Identities.



# Exploring Variations across Race/Ethnicity 

## {.sidebar}
1. How does Perceived Neighborhood Safety vary for individuals with different race/ethnicity? 

Project Description: 
This cross-sectional study collected data on 1329 racially and ethnically diverse adults. a large proportion of the sample also identifies as Sexual Gender Minority (SGM). The goal of the original study is to identify whether adult SGM subgroups differ in terms of their mental and physical health behaviors and risk, as well as factors that affect the health of those with diverse sexual and gender identities. 

Here, we use several data visualizations to explore the data before constructing strata for analyses using Multilevel Analysis of Individual Heterogeneity and Discriminatory Accuracy (MAIHDA).

Data Source: Health Behaviors of Sexually and Gender Diverse Individuals (PIs: Dr. Austin Folger and Dr. Nichole Kelly).

## Plots

```{r}
#| title: "Data Viz 1 - Final"
#| fig-width: 12
#| fig-height: 7
ggplot(df, aes(x = fct_reorder(race_strata, newsa_cri, median, na.rm = TRUE), y = newsa_cri, fill = race_strata)) +
  geom_boxplot(alpha = 0.7) + 
  scale_fill_brewer(palette = "Set1") +
  theme_classic() + 
  coord_flip() +
  labs(x = "", 
       y = "Perceived Neighborhood Safety Score (1-4)", 
       fill = "Race/Ethnicity",
       title = "Perceived Neighborhood Safety from Crime by Racial and Ethnic Group", 
       caption = "Note: Higher scores indicates safer neighborhood perception.")

```


## Plots {.tabset}

```{r}
#| title: "Data Viz 1 (Ver. 1)"

ggplot(df, aes(x = newsa_cri)) +
  geom_bar() + theme_classic() + 
  facet_wrap(~race_strata) +
  coord_flip() +
  labs(x = "Perceived Neighborhood Safety Score (1-4)", y = "N", title = "Distribution of Perceived Neighborhood Safety from Crime by Racial and Ethnic Groups", caption = "Note: Higher Score Indicating Safer Neighborhood")

```


```{r}
#| title: "Data Viz 1 (Ver. 2)"

ggplot(df, aes(x = newsa_cri)) +
  geom_bar() + theme_classic() + 
  facet_wrap(~race_strata) +
  coord_flip() +
  gghighlight(newsa_cri == 4 | newsa_cri == 1) +
  labs(x = "Perceived Neighborhood Safety Score (1-4)", y = "N", title = "Distribution of Perceived Neighborhood Safety from Crime by Racial and Ethnic Groups", caption = "Note: Higher Score Indicating Safer Neighborhood")

```


```{r}
#| title: "Data Viz 1 (Ver. 3)"
#| fig-width: 11
#| fig-height: 6
ggplot(df, aes(x = race_strata, y = newsa_cri)) +
  geom_boxplot(alpha = 0.7) + 
  theme_classic() +
  coord_flip() +
  labs(x = "Racial and Ethnic Groups", y = "Perceived Neighborhood Safety Score (1-4)", title = "Distribution of Perceived Neighborhood Safety from Crime by Racial and Ethnic Group", caption = "Note: Higher Score Indicating Safer Neighborhood")

```


# Exploring Variations across Gender Identities 

## {.sidebar}

2. How does Perceived Neighborhood Safety vary for individuals with different gender identities? How does this vary by gender identities and race/ethnicity?

Project Description: 
This cross-sectional study collected data on 1329 racially and ethnically diverse adults. a large proportion of the sample also identifies as Sexual Gender Minority (SGM). The goal of the original study is to identify whether adult SGM subgroups differ in terms of their mental and physical health behaviors and risk, as well as factors that affect the health of those with diverse sexual and gender identities. 

Here, we use several data visualizations to explore the data before constructing strata for analyses using Multilevel Analysis of Individual Heterogeneity and Discriminatory Accuracy (MAIHDA).

Data Source: Health Behaviors of Sexually and Gender Diverse Individuals (PIs: Dr. Austin Folger and Dr. Nichole Kelly).


## Plots

```{r}
#| title: "Data Viz 2 - Final"
#| fig-width: 12
#| fig-height: 7

ggplot(df, aes(x = gender_coded, y = newsa_cri)) +
  geom_boxplot(aes(group = gender_coded, fill = gender_coded)) + 
  theme_classic() +
  theme(strip.background = element_blank(),
        strip.text = element_text(face = "bold")) +
  scale_fill_brewer(palette = "Set3") +
  facet_wrap(~race_strata) +
  labs(x = "Gender Identity", 
       fill = "Gender Identity", 
       y = "Perceived Neighborhood Safety Score (1-4)", 
       title = "Perceived Neighborhood Safety from Crime by Gender Identity and Race/Ethnicity",
       caption = "Note: Higher scores indicates safer neighborhood perception.")

```


## Plots {.tabset}

```{r}
#| title: "Data Viz 2 (Ver. 1)"

ggplot(df, aes(x = gender_coded, y = newsa_cri)) +
  geom_boxplot() + theme_classic() +
  labs(x = "Gender Identity", y = "Perceived Neighborhood Safety Score (1-4)", title = "Distribution of Perceived Neighborhood Safety from Crime by Gender Identity", caption = "Note: Higher Score Indicating Safer Neighborhood")

```


```{r}
#| title: "Data Viz 2 (Ver. 2)"

ggplot(df, aes(x = gender_coded, y = newsa_cri)) +
  geom_boxplot() + theme_classic() + 
  stat_summary(fun=mean, geom="point", shape=20, size=4, color="red", position=position_dodge(0.75)) +
  stat_summary(fun.data=mean_se, geom="errorbar", width=0.2, color="red", position=position_dodge(0.75)) +
  labs(x = "Gender Identity", y = "Perceived Neighborhood Safety Score (1-4)", title = "Distribution of Perceived Neighborhood Safety from Crime by Gender Identity", caption = "Note: Higher Score Indicating Safer Neighborhood;
       mean/se were denoted with red circle and red lines.")

```


```{r}
#| title: "Data Viz 2 (Ver. 3)"
#| fig-width: 11
#| fig-height: 6

ggplot(df, aes(x = gender_coded, y = newsa_cri)) +
  geom_boxplot() + theme_classic() +
  facet_wrap(~race_strata) +
  labs(x = "Gender Identity", y = "Perceived Neighborhood Safety Score (1-4)", title = "Distribution of Perceived Neighborhood Safety from Crime by Gender Identity and Race/Ethnicity", caption = "Note: Higher Score Indicating Safer Neighborhood")

```


# Exploring Variations across Income Status

## {.sidebar}

3. How does Perceived Neighborhood Safety vary for individuals with different income status? How does this vary by income status, race/ethnicity and gender?

Project Description: 
This cross-sectional study collected data on 1329 racially and ethnically diverse adults. a large proportion of the sample also identifies as Sexual Gender Minority (SGM). The goal of the original study is to identify whether adult SGM subgroups differ in terms of their mental and physical health behaviors and risk, as well as factors that affect the health of those with diverse sexual and gender identities. 

Here, we use several data visualizations to explore the data before constructing strata for analyses using Multilevel Analysis of Individual Heterogeneity and Discriminatory Accuracy (MAIHDA).

Data Source: Health Behaviors of Sexually and Gender Diverse Individuals (PIs: Dr. Austin Folger and Dr. Nichole Kelly).


## Plots

```{r}
#| title: "Data Viz 3 - Final"
#| fig-width: 14
#| fig-height: 8

ggplot(df, aes(x = income, y = gender_coded, 
               label = round(newsa_cri, digits = 0))) +
  geom_tile(aes(fill = newsa_cri)) + 
  facet_wrap(~race_strata) + 
  scale_fill_distiller(palette = "RdBu", direction = 1) +
  labs(x = "Income", y = "Gender Identity", fill = "Perceived Neighborhood Safety Score (1-4)", title = "Perceived Neighborhood Safety from Crime by Income Status, Gender Identity and Race/Ethnicity", caption = "Note: Higher scores indicates safer neighborhood perception.") + theme_classic() +
  theme(strip.background = element_blank(),
        strip.text = element_text(face = "bold"))

```


## Plots {.tabset}

```{r}
#| title: "Data Viz 3 (Ver. 1)"

ggplot(df, aes(x = income, y = newsa_cri)) +
  geom_boxplot() + theme_classic() +
  facet_wrap(~race_strata) +
  labs(x = "Income", y = "Perceived Neighborhood Safety Score (1-4)", title = "Distribution of Perceived Neighborhood Safety from Crime by Income Status and Race/Ethnicity", caption = "Note: Higher Score Indicating Safer Neighborhood")

```


```{r}
#| title: "Data Viz 3 (Ver. 2)"
#| fig-width: 11
#| fig-height: 6


ggplot(df, aes(x = income, y = gender_coded, 
               label = round(newsa_cri, digits = 0))) +
  geom_tile(aes(fill = newsa_cri)) + 
  facet_wrap(~race_strata) + 
  scale_fill_distiller(palette = "RdBu", direction = -1) +
  labs(x = "Income", y = "Gender Identity", fill = "Perceived Neighborhood Safety Score (1-4)", title = "Heatmap of Perceived Neighborhood Safety from Crime by Income Status, Gender Identity and Race/Ethnicity", caption = "Note: Higher Score Indicating Safer Neighborhood") + theme_classic()

```


# Variations across Constructed Strata

## {.sidebar}

4. How does Perceived Neighborhood Safety vary across strata of individuals with intersectional identities (race/ethnicity, gender, income)?

Project Description: 
This cross-sectional study collected data on 1329 racially and ethnically diverse adults. a large proportion of the sample also identifies as Sexual Gender Minority (SGM). The goal of the original study is to identify whether adult SGM subgroups differ in terms of their mental and physical health behaviors and risk, as well as factors that affect the health of those with diverse sexual and gender identities. 

Here, we use several data visualizations to explore the data before constructing strata for analyses using Multilevel Analysis of Individual Heterogeneity and Discriminatory Accuracy (MAIHDA).

Data Source: Health Behaviors of Sexually and Gender Diverse Individuals (PIs: Dr. Austin Folger and Dr. Nichole Kelly).


## Plots

```{r}
#| title: "Data Viz 4 - Final"

df$st_race <- as.factor(df$st_race)
df$st_race <- fct_recode(df$st_race, 
                         "Non-Hispanic White" = "1",
                         "Non-Hispanic POC" = "2",
                         "Hispanic White" = "3",
                         "Hispanic POC" = "4")
df <- rename(df, "Race/Ethnicity" = st_race)

ggplot(df, aes(x = newsa_cri, 
               y = fct_reorder(strata_name, newsa_cri), 
               fill = `Race/Ethnicity`, shape = `Race/Ethnicity`)) +
  stat_summary(fun=mean, geom="point", size=4, position=position_dodge(0.75)) +
  stat_summary(fun.data=mean_se, geom="errorbar", width=0.2, position=position_dodge(0.75)) +
  scale_shape_manual(values = c(21, 22, 23, 24)) +
   scale_fill_brewer(palette = "Set1") +
  theme_classic() +
  labs(x = "Perceived Neighborhood Safety Score (1-4)", 
       y = "Characteristics of Social Strata", 
       title = "Perceived Neighborhood Safety from Crime by Intersectional Strata", caption = "Note: Higher scores indicates safer neighborhood perception; strata were constructed based on individual-reported race/ethnicity, gender identities and income levels.")


```


## Plots {.tabset}

```{r}
#| title: "Data Viz 4 (Ver. 1)"

ggplot(df, aes(x = strata, y = newsa_cri)) +
  geom_boxplot() + theme_classic() +
  coord_flip() +
  labs(x = "Strata IDs", y = "Perceived Neighborhood Safety Score (1-4)", title = "Distribution of Perceived Neighborhood Safety from Crime by Intersectional Strata", caption = "Note: Higher Score Indicating Safer Neighborhood; strata were constructed with race/ethnicity, gender identities and income levels.")


```


```{r}
#| title: "Data Viz 4 (Ver. 2)"

ggplot(df, aes(x = newsa_cri, 
               y = fct_reorder(strata, newsa_cri))) +
  geom_boxplot() + theme_classic() +
  labs(x = "Perceived Neighborhood Safety Score (1-4)", y = "Strata IDs", title = "Distribution of Perceived Neighborhood Safety from Crime by Intersectional Strata", caption = "Note: Higher Score Indicating Safer Neighborhood; strata were constructed with race/ethnicity, gender identities and income levels.")


```


```{r}
#| title: "Data Viz 4 (Ver. 3)"

ggplot(df, aes(x = newsa_cri, 
               y = fct_reorder(strata, newsa_cri))) +
  geom_boxplot() + 
  stat_summary(fun=mean, geom="point", shape=20, size=4, color="red", position=position_dodge(0.75)) +
  stat_summary(fun.data=mean_se, geom="errorbar", width=0.2, color="red", position=position_dodge(0.75)) +
  theme_classic() +
  labs(x = "Perceived Neighborhood Safety Score (1-4)", y = "Strata IDs", title = "Distribution of Perceived Neighborhood Safety from Crime by Intersectional Strata", caption = "Note: Higher Score Indicating Safer Neighborhood; strata were constructed with race/ethnicity, gender identities and income levels.")


```

