---
title: "EDLD 652: Final Project - Chen Su"
format: 
  dashboard:
    theme: vapor
    orientation: columns
output-dir: docs
---

```{r}
library(viridis)
library(RColorBrewer)
library(colorspace)
library(ggrepel)
library(tidyverse)
library(gghighlight)
library(plotly)

df <- readRDS("ms1_cs.rds")

```

Final Project EDLD


# Page 1

## {.sidebar}
Data Source: Health Behaviors of Sexually and Gender Diverse Individuals (PIs: Dr. Austin Folger and Dr. Nichole Kelly)

Project Description: 
This cross-sectional study collected data on 1300 racially and ethnically diverse adults. The goal of the study is to identify whether adult Sexual Gender Minority (SGM) subgroups differ in terms of their mental and physical health behaviors and risk, as well as factors that affect the health of those with diverse sexual and gender identities. 

The data used for all the visualizations here focus on a subset of the main study sample, and only included minoritized racial and ethnic individuals (n = 912). 

## Plots {.tabset}

```{r}
#| title: "Data Viz 1"

ggplot(df, aes(x = age, y = bmi_cleaned, color = gender_coded)) +
  geom_point(size = 2) + theme_classic()

```


```{r}
#| title: "Data Viz 1.1"

ggplot(df, aes(x = age, y = bmi_cleaned, color = gender_coded)) +
  geom_point(size = 2) + theme_classic()

```


# Page 2

## {.sidebar}
Sidebar for page 2.

## Plots {.tabset}

```{r}
#| title: "Data Viz 2"
ggplot(df, aes(x = age, y = bmi_cleaned, color = gender_coded)) +
  geom_point(size = 2) + theme_classic()

```

```{r}
#| title: "Data Viz 2.1"
ggplot(df, aes(x = age, y = bmi_cleaned, color = gender_coded)) +
  geom_point(size = 2) + theme_classic()

```


# Page 3

## {.sidebar}
Sidebar for page 3.

## Plots {.tabset}

```{r}
#| title: "Data Viz 3"
ggplot(df, aes(x = age, y = bmi_cleaned, color = gender_coded)) +
  geom_point(size = 2) + theme_classic()
```


```{r}
#| title: "Data Viz 3.1"
ggplot(df, aes(x = age, y = bmi_cleaned, color = gender_coded)) +
  geom_point(size = 2) + theme_classic()
```


# Page 4

## {.sidebar}
Sidebar for page 4.

## Plots {.tabset}

```{r}
#| title: "Data Viz 4"
p <- ggplot(df, aes(x = age, y = bmi_cleaned, color = gender_coded)) +
  geom_point(size = 2) + theme_classic()

```


```{r}
#| title: "Data Viz 4.1"
p <- ggplot(df, aes(x = age, y = bmi_cleaned, color = gender_coded)) +
  geom_point(size = 2) + theme_classic()

ggplotly(p)
```
